---
title: "Intro To Regression"
author: "Haydt, Donahue, and Worm"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r libraries}
library(lme4)
library(AICcmodavg)
library(GGally)
library(sjPlot)
library(lmerTest)
library(MuMIn)
```

2 Data Sets:
Chesapeake and Ohio Canal Turtle Data
Shrike Habitat Data

## Bring in Data Sets
```{r}
turtle_data <- read.csv("data/CO-turtles.csv")
shrike_data <- read.csv("data/shrike_habitat_data.csv")
```

## Linear Models
# Run Models after Check Data for Normality and Equal Variances
```{r}
# Subset data to get rid of repeat individuals
subset_turtle <- dplyr::filter(turtle_data, recap == "N")

# Does turtle species or carapace length impact individual mass?
lm1 <- lm(data = subset_turtle, mass ~ species) # same as an anova
lm2 <- lm(data = subset_turtle, mass ~ carapace)
lm3 <- lm(data = subset_turtle, mass ~ species + carapace) # examine as independent terms
lm4 <- lm(data = subset_turtle, mass ~ species*carapace) # examine species and carapace as both interaction terms and independent terms
lm5 <- lm(data = subset_turtle, mass ~ species:carapace) # just looking at interactions

summary(lm1)
summary(lm2)
summary(lm3)
summary(lm4)
summary(lm5)
```

## Generalized Linear Models
```{r}
# Subset data to get rid of repeat individuals
# subset_turtle <- dplyr::filter(turtle_data, recap == "N")

# Does number of leeches observed differ by species or site? 
glm1 <- glm(data = subset_turtle, leeches ~ species,
            family = poisson)
glm2 <- glm(data = subset_turtle, leeches ~ species*mass,
            family = poisson)
glm3 <- glm(data = subset_turtle, leeches ~ species*mass + site,
            family = poisson)
summary(glm1)
summary(glm2)
summary(glm3)
```


### MIXED MODELS

## Linear Mixed Models 
```{r}
# Does carapace length impact individual mass?
lmm1 <- lmer(data = turtle_data, mass ~ species*carapace + (1|uniqueID))
lmm2 <- lmer(data = turtle_data, mass ~ carapace + (1|uniqueID/species))
summary(lmm1)
summary(lmm2)
```

## Generalized Linear Mixed Models - Can used distributions such as POISSON (for count data) or BINOMIAL (for binomial data, 0s and 1s)
```{r}
# Turtles: Does the species impact the number of leeches observed?
glmm1 <- glmer(data = turtle_data, 
               leeches ~ species 
               + (1|site), family = poisson)
summary(glmm1)

library(lmerTest) # If you are using lmer model and can't see the p-value output
```


X_________________________________________________________________________
     Shrike Data Example -- Generalized Linear Mixed Model Selection
X_________________________________________________________________________


##___Site-Level Habitat Characteristics Model Selection__

# Check for correlation between predictor variables
```{r}

## Site-Level Characteristics

# (Values of r>0.7 typically considered highly correlated, decide which to remove)

cor(shrike_data[, c("PerchH", "Crops11", "Water11")], use = "complete.obs")

ggpairs(shrike_data, columns = c("PerchH", "Crops11", "Water11"))
```

## Build and compare models for site-level habitat characteristics
```{r}
# Build full model with all variables of interest

site_full<- glmer(Use ~ PerchH + Crops11 + Water11 + 
                    (1|PointID) + (1|BirdID), family = binomial, 
                    data = shrike_data)

# Singularity error may result if random effect explains too little variance,
# But random errors are important to include here to pair used/random points and account for psuedo-replication

# Run models with all possible combinations of predictor variables and null with dredge function

options(na.action = "na.fail") #Ensure each model is run with the same dataset and no rows are excluded because of NA values
model_set1 <- dredge(site_full)

# Analyze models by comparing AICc values
# Models in table with deltaAICc (delta) values < 2 considered equivalent
print(model_set1)
```

## Build top models for site-level characteristics
```{r}
#Build top models (delta AICc<2) from dredge table output --> print(model_set1)

site_top1<- glmer(Use ~ PerchH + Water11 + 
                 (1|PointID), family = binomial, data = shrike_data)

site_top2<- glmer(Use ~ Crops11 + PerchH + Water11 + 
                 (1|PointID) + (1|BirdID), family = binomial, nAGQ =1, 
                 data = shrike_data)

# Evaluate confidence intervals to evaluate which parameters may be important in explaining response variable
# Variable is important if its 85% confidence interval does not overlap (see Arnold 2010)

confint(site_top1,level=0.85, method="Wald")
confint(site_top2,level=0.85, method="Wald")

# Use summary function to get beta estimates for each important predictor variable
#Beta > 0 --> positive correlation, Beta < 0 --> negative correlation

summary(site_top1)
summary(site_top2)
```

## Plot Top Models for Site-level habitat characteristics
```{r}
#Make plot with predicted values to visualize model correlation trends

plot_model(site_top1, type = "pred", terms = "Water11")
plot_model(site_top1, type = "pred", terms="PerchH [all]") #[all ensures smooth plot]


# Make boxplot with predicted values to visualize model correlation trends for binary variables

#Get predicted probabilities from the model (fine_top1)
predicted_values <- predict(site_top1, type = "response")

# Create a new data frame that includes the predicted and actual values for Water11
prediction_data <- data.frame(
  Water11 = shrike_data$Water11,
  Predicted = predicted_values
)

# Create a boxplot with ggplot2
ggplot(prediction_data, aes(x = factor(Water11), y = Predicted)) +
  geom_boxplot(fill = "lightblue", color = "blue") +  # Boxplot with color customization
  geom_jitter(width = 0.2, alpha = 0.5) +             # Add jittered points for better visualization
  labs(title = "Predicted Probabilities by Water11", 
       x = "Water11", y = "Predicted Probability") +
  theme_minimal()
```
X__________________________________________________________________


##___Fine-Scale Habitat Characteristics Model Selection__

# Check for correlation between predictor variables
```{r}

# Rescale Wire Length since we have much larger values (i.e. 1000 m) when compared to other variables
rescale<-function(x) (x-min(x)) / (max(x) - min(x)) * 100
Wire100scale<-rescale(shrike_data$Wire100)

# Check for correlation of variables to be included in models
# (Values of r>0.7 typically considered highly correlated, decide which to remove)

cor(shrike_data[, c("Ditches100", "GrassWavg100", "GrassHavg100", "Wire100")], use = "complete.obs")

ggpairs(shrike_data, columns = c("Ditches100", "GrassWavg100", "GrassHavg100", "Wire100"))
```

## Build and compare models for fine-scale habitat characteristics
```{r}
# Build full model with all variables of interest

fine_full<- glmer(Use ~ Ditches100 + GrassWavg100 + 
                    GrassHavg100 + Wire100scale + 
                    (1|PointID) + (1|BirdID), 
                  family = binomial, nAGQ =1, data = shrike_data)

# Run models with all possible combinations of predictor variables and null with dredge function

options(na.action = "na.fail") #Ensure each model is run with the same dataset and no rows are excluded because of NA values

model_set2 <- dredge(fine_full)

# Analyze models by comparing AICc values
# Models in table with deltaAICc (delta) values < 2 considered equivalent
print(model_set2)

# Build top models (delta AICc<2) from dredge table output --> print(model_set1)

fine_top1<- glmer(Use ~ GrassHavg100 +
                    (1|PointID) + (1|BirdID), family = binomial, 
                  data = shrike_data)

fine_top2<- glmer(Use ~ Ditches100 + GrassHavg100 + Wire100scale +
                    (1|PointID) + (1|BirdID), family = binomial, nAGQ =1, 
                  data = shrike_data)

fine_top3<- glmer(Use ~ GrassWavg100 + GrassHavg100 + Wire100scale +
                    (1|PointID) + (1|BirdID), family = binomial, nAGQ =1,
                  data = shrike_data)

# Evaluate confidence intervals to evaluate which parameters may be important in explaining response variable
# Variable is important if its 85% confidence interval does not overlap (see Arnold 2010)

confint(fine_top1,level=0.85, method="Wald")
confint(fine_top2,level=0.85, method="Wald")
confint(fine_top3,level=0.85, method="Wald")

# Use summary function to get beta estimates for each important predictor variable
# Beta > 0 --> positive correlation, Beta < 0 --> negative correlation

summary(fine_top1)
summary(fine_top2)
summary(fine_top3)
```

##Plot Tops Models for Fine-scale habitat characteristics

```{r}
# Model plots to visualize correlation trends

plot_model(fine_top1, type = "pred", terms = "GrassHavg100 [all]")
plot_model(fine_top2, type = "pred", terms="Wire100scale")
```
X_________________________________________________________________


##___Broad-Scale Habitat Characteristics Model Selection__

# Check for correlation between predictor variables
```{r}
#(Values of r>0.7 typically considered highly correlated, decide which to remove)

cor(shrike_data[, c("Cotton500", "Rice500", "Soy500", "Develop500", "Forest500")], use = "complete.obs")

ggpairs(shrike_data, columns = c("Cotton500", "Rice500", "Soy500", "Develop500", "Forest500"))
```

## Build and compare models for broad-scale habitat characteristics
```{r}
#Build full model with all variables of interest

broad_full<- glmer(Use ~ Cotton500 + Rice500 + Soy500 + 
                     Develop500 + Forest500 +
                    (1|PointID) + (1|BirdID), 
                   family = binomial, nAGQ =1, 
                   data = shrike_data)

# Run models with all possible combinations of predictor variables and null with dredge function

options(na.action = "na.fail") #Ensure each model is run with the same dataset and no rows are excluded because of NA values

model_set3 <- dredge(broad_full)

# Analyze models by comparing AICc values
# Models in table with deltaAICc (delta) values < 2 considered equivalent
print(model_set3)
```

## Build Top Models for Broad-scale habitat characteristics
```{r}
# Build top models (delta AICc<2) from dredge table output --> print(model_set1)

broad_top1<- glmer(Use ~ Develop500 +
                    (1|PointID) + (1|BirdID), 
                   family = binomial, data = shrike_data)

broad_top2<- glmer(Use ~ Develop500 + Soy500 +
                    (1|PointID) + (1|BirdID), family = binomial, nAGQ =1, 
                   data = shrike_data)

broad_top3<- glmer(Use ~ Develop500 + Forest500 +
                    (1|PointID) + (1|BirdID), 
                   family = binomial, nAGQ =1, data = shrike_data)

broad_top4<- glmer(Use ~ Develop500 + Rice500 +
                     (1|PointID) + (1|BirdID), family = binomial, 
                   nAGQ =1, data = shrike_data)

broad_top5<- glmer(Use ~ Develop500 + Cotton500 +
                     (1|PointID) + (1|BirdID), family = binomial, 
                   nAGQ =1, data = shrike_data)


# Evaluate confidence intervals to evaluate which parameters may be important in explaining response variable
# Variable is important if its 85% confidence interval does not overlap (see Arnold 2010)

confint(broad_top1,level=0.85, method="Wald")
confint(broad_top2,level=0.85, method="Wald")
confint(broad_top3,level=0.85, method="Wald")
confint(broad_top4,level=0.85, method="Wald")
confint(broad_top5,level=0.85, method="Wald")

# Use summary function to get beta estimates for each important predictor variable
# Beta > 0 --> positive correlation, Beta < 0 --> negative correlation

summary(broad_top1)
```

## Plot Top Models for Broad-scale habitat characteristics
```{r}
#Model plots to visualize correlation trends

plot_model(broad_top1, type = "pred", terms = "Develop500 [all]")
```














