---
title: "Intro To Regression"
author: "Haydt, Donahue, and Worm"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r libraries}
library(lme4)
```

Data Sets:
Shrike Habitat Data - 
Chesapeake and Ohio Canal Turtle Data - 

## Bring in Data Sets
```{r}
turtle_data <- read.csv("data/CO-turtles.csv")
shrike_data <- 
```

## Linear Models
```{r}
# Does turtle species or carapace length impact individual mass?
m1 <- lm(data = turtle_data, mass ~ species) # same as an anova
m2 <- lm(data = turtle_data, mass ~ carapace)
m3 <- lm(data = turtle_data, mass ~ species + carapace) # examine as independent terms
m4 <- lm(data = turtle_data, mass ~ species*carapace) # examine species and carapace as both interaction terms and independent terms
m5 <- lm(data = turtle_data, mass ~ species:carapace) # just looking at interactions

summary(m1)
summary(m2)
summary(m3)
summary(m4)
summary(m5)
```

## Generalized Linear Models
```{r}

```

### MIXED MODELS
Fixed vs. random effects:


## Linear Mixed Models 
```{r}
# Does carapace length impact individual mass?
m1 <- lmer(data = turtle_data, mass ~ carapace + (1|species))
summary(m1)

library(lmerTest)
m2 <- lmerTest::lmer(data = turtle_data, mass ~ carapace + (1|species))
summary(m2)
```

## Generalized Linear Mixed Models - Can use non-normal distributions such as POISSON and BINOMIAL
```{r}
# Does the species impact the number of leeches observed?
m1 <- lmer(data = turtle_data, leeches ~ species + (1|site))
summary(m1)

library(lmerTest)
m2 <- lmerTest::lmer(data = turtle_data, leeches ~ species + (1|site))
summary(m2)
```

## Making Tables with Regression Outputs
```{r}

```

## Viewing Effects (Including interaction effects)
```{r}

```


